### 2024-10-31 letter_to_editor.tex
cover letter sketched

\documentclass[a4paper]{article}

\usepackage[margin=1in]{geometry}

\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\hypersetup{
	unicode,
}

\newcommand{\submission}{X}
\newcommand{\journal}{Y}

\begin{document}

\noindent Dear Editor,\\

We would like to submit our draft ``\textbf{\submission}'' to \textbf{\journal}.\\

Benchmarking is an integral part of computational biology and bioinformatics. Yet benchmarks are far from being standardized, reusable or extensible, and their implementation is cumbersome in terms of repetitive tasks, such as managing reproducible software environments or formalizing computational workflows.\\

In this manuscript we describe the desired properties of benchmarking systems, e.g., software solutions to facilitate, standardize, coordinate and help distributing benchmarks.\\

We think our paper matches \journal's call for .\\

Thank you in advance for processing our submission. \\

Best regards,\\

Izaskun Mallona\\
\indent University of Zurich\\

\vspace{1cm}

\textit{Suggested reviewers}

\begin{itemize}
    \item Dr Salvador Capella Gutierrez, \texttt{salvador.capella@bsc.es}, leader of the Spanish National Bioinformatics (INB) Coordination Node at the Barcelona Supercomputing Center (BSC).
    \item Dr Malte Luecken, \texttt{malte.luecken@helmholtz-muenchen.de}, group leader of the Computational Cellular Health Group, Helmholtz Munich.
\end{itemize}

\end{document}