### 2025-01-31 review_cycle_1/letter_to_editor.tex
Add figure draft

\documentclass[a4paper]{article}

\usepackage[margin=1in]{geometry}

\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\hypersetup{
	unicode,
}

\newcommand{\submission}{Building a continuous benchmarking ecosystem in bioinformatics}
\newcommand{\journal}{PLoS Computational Biology}
\newcommand{\submitter}{Izaskun Mallona and Mark D. Robinson}
\newcommand{\editor}{Dr BF Francis Ouellette, Dr Patricia M Palagi}

\begin{document}

\noindent Dear \editor,\\

Thank you for processing our manuscript \texttt{PCOMPBIOL-D-24-02003}.

\subsection*{Reviewer 1}

\end{document}