### 2025-04-23 0_original_submission/letter_to_editor.tex
before generating the bbl

\documentclass[a4paper]{article}

\usepackage[margin=1in]{geometry}

\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\hypersetup{
	unicode,
}

\newcommand{\submission}{Building a continuous benchmarking ecosystem in bioinformatics}
\newcommand{\journal}{NAR Genomics and Bioinformatics}
\newcommand{\submitter}{Izaskun Mallona and Mark D. Robinson}
\newcommand{\editor}{Dr Ana Rojas}

\begin{document}

Dear \editor,\\

We would like to submit our draft ``\textbf{\submission}'' to \textbf{\journal}.\\

Benchmarking is an integral part of computational biology and bioinformatics. Yet benchmarks are far from being standardized, reusable or extensible, and their implementation is cumbersome in terms of repetitive tasks, such as managing reproducible software environments or formalizing computational workflows.\\

In this manuscript we argue benchmarking can benefit from benchmarking systems, e.g., software solutions to facilitate, standardize, coordinate and help distributing benchmarks, and discuss the desired properties of such benchmarking systems.\\

We think our paper matches \journal's scope and interest in robust and reproducible benchmarking of genomics and bioinformatics methods.\\

Our manuscript is available as a preprint at \url{https://arxiv.org/abs/2409.15472}.\\

Thank you in advance for processing our submission. \\

Best regards,\\

\submitter\\
\indent University of Zurich\\

\vspace{1cm}

\textit{Suggested reviewers}

\begin{itemize}
    \item Dr Salvador Capella Gutierrez, \texttt{salvador.capella@bsc.es}, leader of the Spanish National Bioinformatics (INB) Coordination Node at the Barcelona Supercomputing Center (BSC).
    \item Dr Malte Luecken, \texttt{malte.luecken@helmholtz-muenchen.de}, leader of the Computational Cellular Health Group, Helmholtz Munich.
    \item Prof Dr Carole Goble , \texttt{carole.goble@manchester.ac.uk}, full professor and co-leader of the Information Management Group, School of Computer Science in the University of Manchester.
    \item Prof Dr Johannes Koester, \texttt{johannes.koester@uni-due.de}, professor for Bioinformatics and Computational Oncology, Institute for AI in Medicine (IKIM), University Medicine Essen.
    \item Prof Dr Anne-Laure Boulesteix, \texttt{boulesteix@ibe.med.uni-muenchen.de}, leader of the Biometry in Molecular Medicine group, LMU Munich.
    \item Prof Dr. Serguei Mangul, \texttt{serghei.mangul@gmail.com}, assistant professor of clinical pharmacy, University of Southern California.
    \item Prof Dr. Jean Yang, \texttt{jean.yang@sydney.edu.au}, professor, School of Mathematics and Statistics of the University of Sydney.
    \item Prof. Dr. Gregory Grann, \texttt{ggrant@pennmedicine.upenn.edu}, associate professor of genetics, Perelman School of Medicine
\end{itemize}

\end{document}